\chapter*{Introduction}
The theoretical study of atomic nuclei provides a bridge between nuclear physics and nuclear engineering. Starting from a framework consistent with quantum mechanics, the strong interaction, and its underlying symmetries, modern nuclear theory aims to construct models characterised by a limited number of free parameters and capable of predicting both nuclear structure and reactions across a wide range of systems. While experimental data have long provided invaluable insight into nuclear properties and processes, only a coherent theoretical description allows for systematic extrapolations towards regions of the nuclear chart or physical conditions that remain beyond current experimental reach, thus playing an essential role in applications relevant to nuclear engineering.

In particular, nuclear fission, despite its massive importance in nuclear engineering, remains only partially understood from a microscopic standpoint. Current models that use empirical approaches \cite{MyersSwiatecki1996_NPA,Moller2016_ADNDT}, successfully reproduce global quantities like fission barrier heights, fragment mass distributions, and average neutron multiplicities for well studied nuclei. However, these models may rely on a huge number of parameters, which limit their predictive power when extrapolated to systems which are less investigated experimentally. A fully microscopic understanding of the collective dynamics leading from the compound nucleus to scission, the treatment of quantum many-body correlations \cite{Schunck2016}, and the description of fragment excitation and emission remain among the major open challenges, particularly relevant for the simulation of next-generation reactors, which require the accurate description of nuclei and fuel materials -- far less explored than those employed in traditional thermal systems -- to be correctly predicted. 

In this regard, the approach to the microscopic description of nuclei, is the one of the many-body theory, which starting from the interacting nucleons, aims at building a complete description of the nucleus. The use of phenomenological potentials based on the Woods-Saxon one is still relevant, thanks to its computational feasibility and its capability to include shell effects in a simple manner, but it cannot account for many-body effects.
At the moment, there are two competing frameworks that try to tackle the microscopic description of nuclei,
\begin{enumerate}[i]
\item the \textit{ab-initio} approach \cite{Hagen2014}, where the interaction is in principle exact, derived from controlled approximations of quantum chromodynamics; and
\item the use of effective interactions and nuclear Density Functional Theory\cite{NDFT,Bender2003}.
\end{enumerate}
Ab-initio methods, while technically speaking more rigorous, are still limited as of now, since they can only account for light nuclei or medium-heavy nuclei that can be considered as spherical.
Energy density functionals and effective interactions, such as the Skyrme force, on the other hand are more flexible and less computationally expensive, enabling a much wider representation of nuclei across the whole chart, including heavy nuclei and processes such as fission, fusion, reactions and decays, which are of crucial importance in nuclear engineering.

D Vautherin and D M Brink laid the foundations of the nuclear Hartree-Fock theory using the Skyrme interaction in 1972 \cite{VauhBrinkOriginal}, through spherically symmetric calculations, which are unable to account for nuclear deformations, essential for nuclei far from magic numbers.
Over the years, thanks to the increase in computational performance of modern hardware, codes that are able to represent more coordinates have been written, mainly using basis expansions on the harmonic oscillator \cite{hfodd,Marevic2022_CPC}, which have the downside of not being able to account for near drip line nuclei, due to the diffent asymptotic behaviour of the Gaussian basis in the harmonic oscillator and quasi-resonant states. 

In the past twenty years, the use of Cartesian meshes to better account for such extremal cases has been introduced \cite{oldEv8,Ryssens2016_MOCCa,hfbftt,Maruhn2014_Sky3D}, often times assuming certain approximations, such as plane reflection symmetries \cite{Ryssens2015_EV8} and axial symmetry \cite{Maruhn2014_Sky3D}. The use of fully unconstrained Hartree-Fock methods, of critical importance for exotic deformations, is still a novel endeavour that only a handful of implementations have tackled, due to the high computational cost.

The aim of this work is to explore a new computational approach, the Generalised Conjugate Gradient method, to efficiently solve spatially unconstrained Skyrme functionals. This thesis is organised as follows:
\begin{itemize}
    \item In chapter \ref{chap:intro}, a short, comprehensive introduction to nuclear physics and its open problems is given, as to prime the reader on the essential physical properties of atomic nuclei, starting from phenomenological facts and empirical models. A formal description of nuclear deformations and fission is also given, to highlight the importance of symmetry breaking.
    \item In chapter \ref{chap:methods}, the motivations and objectives of this work are presented, along with a summary of the methods used to achieve them.
    \item In chapter \ref{chap:hf}, the theoretical framework used in the present work is reviewed, by introducing aspects of Hartree-Fock theory, Density Functional Theory and the effective interaction used in this work.
    \item In chapter \ref{chap:numerical}, the numerical methods used in this work are presented, along with actual implementations in writing the code.
    \item In chapter \ref{chap:res_sph}, results for the spherically symmetric case are presented as a way of benchmarking the new implementation of this thesis, along with a description of the main physical quantities we compare.
    \item In chapter \ref{chap:res_def}, benchmarks for the deformed nucleus $^{24}$Mg are shown, after which novel results regarding... are presented.
\end{itemize}
\chapter{Nuclear structure and deformations}
\label{chap:intro}
In this chapter, a very concise introduction to nuclear structure physics is provided, with the goal of delivering the means to understand the essential phenomenology, physical quantities, and terminology of atomic nuclei. In this way, the reader will have the necessary background to understand nuclear deformations and other properties discussed in later sections. The reader may refer to introductory textbooks on nuclear physics, such as Refs. \cite{bertulani2007nuclear,obertelli2021modern} for a more in-depth treatment of the subject and Ref. \cite{ring2004nuclear} for a focus on many-body theory.

This chapter is organised as follows. First, in section \ref{sec:models}, we will review the main empirical facts about nuclides, such as the shapes of particle density distributions, the nuclear binding energies and the phenomenological models employed to describe them. We shall then move on to more advanced topics, that are able to complete the general description of nuclear structure, like nuclear pairing and nuclear deformations in section \ref{sec:pairing_intro} and section \ref{sec:deformations} respectively.
Lastly, in section \ref{sec:fission}, we will review the nuclear fission process, discuss a simple model to describe it, and highlight the importance of deformations to accurately describe it.

\input{chapters/introduction/nuclearmodels.tex}
\input{chapters/introduction/pairing.tex}
\input{chapters/introduction/deformations.tex}
\input{chapters/introduction/fission.tex}